---
title: "metric analysis"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
    toc_float: yes
    number_sections: true
    code_folding: hide
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(1234)

library(dplyr)
library(stringr)
library(readr)
resistance <- read_csv("./ED-outputs/results/metric/resistance_values.csv")
resilience <- read_csv("./ED-outputs/results/metric/resilience_values.csv")
met <- read_csv("./ED-outputs/results/constant_annual_met_data.csv")

# The variable table, contains inforamtion about the met variable abreviation with a descrioption. 
var_info <- data.table::data.table(met_var = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp', 'GPP', 'grd'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature', 
                                       'pre-dist. productivity', 'total wind'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-1', 'kgH2O kgAir-1', 'K', 'MgC ha-1 year-1', 'm s-1'), stringsAsFactors=FALSE)




# Add 2000 GPP to the met df since it is a "condition". 
ED_output <- read.csv("ED-outputs/results/exp-constant-yr.csv", stringsAsFactors = FALSE) %>% 
  filter(scn == "harvest_0" & variable == "GPP" & year == 2000) %>% 
  select(met, value, variable, value, units = unit)

met %>% 
  filter(variable %in% c("ugrd", "vgrd")) %>% 
  mutate(value = value^2) %>% 
  group_by(year, met) %>% 
  summarise(value = sqrt(sum(value))) %>% 
  ungroup() %>%
  mutate(units = "m s-1", 
         description = "total winnd", 
         variable = "grd") -> 
  wind_df
  
met %>% 
  filter(!variable %in% c("ugrd", "vgrd")) %>% 
  bind_rows(wind_df, 
            ED_output) ->
  met

resistance %>% 
  select(scn, variable, met, value = trough_resistance, severity) %>% 
  mutate(metric = "resistance") ->
  combined
resilience %>% 
  select(scn, variable, met, value = resilience, severity) %>% 
  mutate(metric = "resilience") %>% 
  bind_rows(combined) ->
  combined
```










# Met data

Take a look at the distribution of met data, and perform a Principal Components
Analysis (PCA) to examine auto-correlation between the various met variables.

```{r, echo=FALSE}
des_order <- c("Air Temperature\n(K)", "Visible Beam Downward Solar Radiation\n(W m-2)",
               "Atmospheric Pressure\n(Pa)", "Visible Diffuse Downward Solar Radiation\n(W m-2)", 
               "Downward Long Wave Radiation\n(W m-2)", "Specific Humidity\n(kgH2O kgAir-1)", 
                "Precipitation Rate\n(kgH2O m-2 s-1)", "Total Wind\n(m s-1)", 
               "Pre-Dist. Productivity\n(MgC ha-1 year-1)")       


met %>% 
  # hgt is constant; remove
  filter(variable != "hgt") %>%   
  group_by(met, variable) %>% 
  summarise(value = mean(value), .groups = "drop") ->
  met2


met2 %>% 
  mutate(met_var = variable) %>% 
  inner_join(var_info, by = "met_var") %>%  
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
  filter(description %in% des_order) -> 
  to_plot_met2

met2 %>% 
  filter(met == 2019) -> 
  met2019

met2019 %>% 
  mutate(met_var = variable) %>% 
  left_join(var_info, by = "met_var") %>%  
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
  to_plot_2019


to_plot_met2$description <- factor(to_plot_met2$description, des_order, 
                              ordered = TRUE) 
to_plot_2019$description <- factor(to_plot_2019$description, des_order, 
                              ordered = TRUE) 

library(ggplot2)
theme_set(theme_minimal())
ggplot(to_plot_met2, aes(met, value)) +
  geom_point() + 
  #geom_point(data = met2019, aes(met, value), color = "#D55E00") +
  facet_wrap(~description, scales = "free_y", labeller = label_wrap_gen(width = 20)) +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = NULL, y = 'Mean annual conditions value') -> 
  plot 

ggsave(plot, filename = "./C.analysis/figures/conditions_yr.png", width = 10, height = 6)
plot

library(tidyr)
met2_wide <- pivot_wider(met2, id_cols = "met", names_from = "variable")
pca_met <- prcomp(met2_wide[-1], scale. = TRUE)  # -1 removes met year column 
bit(pca_met, scale = 0)
pairs(met2_wide[-1])
```

# Metric data

```{r}
ggplot(resistance, aes(trough_resistance, fill = severity)) +
  geom_histogram(bins = 20) + 
  ggtitle("Resistance") +
  facet_wrap(~variable)

ggplot(resilience, aes(resilience, fill = severity)) +
  geom_histogram(bins = 20) + 
  ggtitle("Resilience") +
  facet_wrap(~variable)
```

# Met-metric data {.tabset}

We see from the met analysis above that vbdsf and vddsf are highly
correlated with nbdsf and nddsf respectively; remove one pair.

Take a first look at the relationship between output variables and met variables.

## Resistance

```{r, fig.height=8}
# Build a combined met-metric dataset
combined %>% 
  select(metric, variable, met, value) %>% 
  # We see from the met analysis above that vbdsf and vddsf are highly
  # correlated with nbdsf and nddsf respectively; remove one pair
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  metric_combined

resistance %>% 
  select(variable, met, trough_resistance) %>% 
  left_join(select(met2_wide, -nbdsf, -nddsf), by = "met") ->
  resist_combined

metric_combined %>% 
  pivot_longer(dlwrf:vddsf, values_to = "met_value") ->
  metric_comb_long

p_mm <- ggplot(filter(metric_comb_long, metric == "resistance"), 
               aes(met_value, value, color = variable)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, color = "black", linetype = 5) +
  facet_wrap(~ variable + name, scales = "free") + 
  ggtitle("Resistance") +
  theme(legend.position = "none", axis.text = element_text(size = 5))
print(p_mm)
```

## Resilience

```{r, fig.height=8}
p_mm <- p_mm %+% filter(metric_comb_long, metric == "resilience")
print(p_mm + ggtitle("Resilience"))
```

# Variable importance {.tabset}

Fit a Random Forest model to each output variable (GPP, etc.) and then calculate
variable importance for each model. Make partial dependence plots for the
three most important met variables for each output variable.

```{r, message=FALSE}
# Fit a Random Forest for each variable (across all treatment severities)
# Note the r.f. functions don't play well with tibbles, so change to data frame
library(randomForest)
d <- as.data.frame(metric_combined)
rc_all <- split(d, interaction(d$metric, d$variable))
rf_all <- lapply(rc_all, function(x) 
  # remove metric, variable, met columns and run RF
  randomForest(value ~ ., data = x[-1:-3], importance = TRUE)
)

# Importance metrics and partial plot data
import <- list()
partial <- list()
for(var in names(rf_all)) {
  mod <- rf_all[[var]]
  metric <- rc_all[[var]]$metric[1]
  variable <- rc_all[[var]]$variable[1]
  
  imp <- as.data.frame(importance(mod))
  imp$met_var <- rownames(imp)
  imp$metric <- metric
  imp$variable <- variable
  import[[var]] <- as_tibble(imp)
  
  impvar <- rownames(imp)[order(imp[, 1], decreasing = TRUE)]
  for (i in seq_along(impvar[1:3])) {
    pp <- partialPlot(mod, rc_all[[var]], impvar[i], plot = FALSE)
    partial[[paste(var, i)]] <- tibble(metric = metric,
                                       variable = variable,
                                       met_var = impvar[i],
                                       x = pp$x, y = pp$y)
  }
}

import <- bind_rows(import)
partial <- bind_rows(partial)
```

## Resistance

```{r}


library(forcats)
f_imp_plot <- function(x) {
  ggplot(x, aes(fct_reorder(met_var, `%IncMSE`), `%IncMSE`)) + 
    geom_col() +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
    coord_flip() +
    xlab(NULL) +
    ggtitle(unique(x$variable)) 
}

import %>%
  left_join(var_info) %>% 
  mutate(met_var = str_to_title(description)) -> 
  import2


d <- filter(import2, metric == "resistance")
xx <- lapply(split(d, d$variable), f_imp_plot)

library(cowplot)
varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
print(varimp_plot)

ggsave(varimp_plot, filename = "./C.analysis/figures/varimp_plot_resistance.png", width = 10, height = 6)

save_plot("varimp_resistance.png", varimp_plot)

p_pdp <- ggplot(filter(partial, metric == "resistance"), 
                aes(x, y, color = variable)) + 
  geom_line() + 
  xlab("Met var value") + ylab("Resistance") +
  ggtitle("Partial dependence plot") +
  facet_grid(variable ~ met_var, scales = "free")
print(p_pdp)
```

## Resilience

```{r}
# Re-do above plots but now for resilience
d <- filter(import2, metric == "resilience")
xx <- lapply(split(d, d$variable), f_imp_plot)
varimp_plot <- plot_grid(xx[["GPP"]], xx[["NEP"]], xx[["NPP"]], xx[["Rh"]])
print(varimp_plot)
ggsave(varimp_plot, filename = "./C.analysis/figures/varimp_plot_resilience.png", width = 10, height = 6)

save_plot("varimp_resilience.png", varimp_plot)

p_pdp <- p_pdp %+% filter(partial, metric == "resilience")
print(p_pdp + ylab("Resilience"))
```

# Summary

```{r}

var_info <- data.table::data.table(met_var = c('dlwrf', 'nbdsf', 'nddsf', 'vbdsf',
                                    'vddsf', 'prate', 'pres', 'hgt', 
                                    'ugrd', 'vgrd', 'sh', 'tmp', 'GPP'), 
                       description = c('downward long wave radiation', 
                                       'near infrared beam downward solar radiation', 
                                       'near IR diffuse downward solar radiation', 
                                       'visible beam downward solar radiation',
                                       'visible diffuse downward solar radiation', 
                                       'precipitation rate', 
                                       'atmospheric pressure', 
                                       'geopotential height', 
                                       'zonal wind',
                                       'meridional wind', 
                                       'specific humidity',
                                       'air temperature', 
                                       'pre-dist. productivity'),  
                       units = c('W m-2', 'W m-2', 'W m-2', 'W m-2',
                                 'W m-2', 'kgH2O m-2 s-1', 'Pa', 'm',
                                 'm s-1', 'm s-', 'kgH2O kgAir-1', 'K', 'MgC ha-1 year-1'), stringsAsFactors=FALSE)



var_info$met_var <- factor(var_info$met_var, unique(var_info$met_var), ordered = TRUE)

partial %>% 
  left_join(var_info) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) -> 
  data_partial


# Combined partial dependence plot
data_partial %>% 
  mutate(metric = if_else(metric == "resistance", "Resistance", "Resilience")) ->  
  to_plot 


des_order <- c("Air Temperature\n(K)", "Visible Beam Downward Solar Radiation\n(W m-2)",
               "Atmospheric Pressure\n(Pa)", "Visible Diffuse Downward Solar Radiation\n(W m-2)", 
               "Downward Long Wave Radiation\n(W m-2)", "Specific Humidity\n(kgH2O kgAir-1)", 
                "Precipitation Rate\n(kgH2O m-2 s-1)", "Pre-Dist. Productivity\n(MgC ha-1 year-1)")       

to_plot$description <- factor(to_plot$description, des_order, 
                              ordered = TRUE)

to_plot %>% 
  ggplot(aes(x, y, color = variable)) + 
  geom_line() + 
  xlab("Conditions variable value") + ylab("Metric value") +
  scale_color_discrete("") +
  #facet_grid(metric ~ description, scales = "free", switch = "y") +
  facet_grid(metric ~ description, 
             labeller = label_wrap_gen(width = 10), scales = "free", switch = "y") +
  theme_bw()+ 
  theme(axis.text.x = element_text(angle = 90)) -> 
  plot 

plot

ggsave(plot, filename = "./C.analysis/figures/cond_metric.png", width = 10, height = 6)
```

For **resistance**, `tmp` (air temperature) and `vbdsf` (sunlight) are among the
top-three most important drivers for all four output variables 
(GPP, NEP, NPP, Rh). Higher temperature and more light translate into higher
resistance for all variables.

This is largely true for **resilience** as well, although
`prate` (precipitation) is important for both NPP and Rh; in fact precipitation
is far and away the most important variable for Rh. Here, however,
higher temperature and more light mean _less_ resilience.

Interesting that when GPP is included as a condition that could explain the resistance & resilience metrics it really only affect the Rh values.    



```{r}
met %>% 
  filter(variable == "GPP") %>% 
  select(met, pre_GPP = value) -> 
  pre_dist_prod

met %>% 
  filter(!variable %in% c("GPP", "hgt")) %>% 
  mutate(description = paste0(str_to_title(description), '\n(', units, ')')) %>% 
  select(met, value, description) -> 
  met_values

equal_breaks <- function(n = 3, s = 0.05, ...){
  function(x){
    # rescaling
    d <- s * diff(range(x)) / (1+2*s)
    signif(seq(min(x)+d, max(x)-d, length=n), digits = 3)
  }
}

met_values %>% 
  left_join(pre_dist_prod) %>% 
  ggplot(aes(value, pre_GPP)) + 
  geom_point() +
  facet_wrap(~description, scales = "free_x", labeller = label_wrap_gen(width = 20)) + 
  labs(y = "Pre-disturbance productivity (MgC ha-1)", 
       x = "Meteorological value") + 
  scale_x_continuous(breaks=equal_breaks(n=3, s=0.05), 
                     expand = c(0.05, 0)) + 
    theme(axis.text.x = element_text(angle=90)) -> 
  plot

ggsave(plot, filename = "./C.analysis/figures/gpp_met.png", width = 10, height = 6)

  

```
